Predicting Spotify Audio Features from Last.fm Tags

### 2023-07-07 paper/main.tex
add intro remainder part

The metadata associated to a music piece is available in multiple formats.
Editorial information or lyrics, for example, are mostly available in text format.
The ability to process images or videos is also required, for example, for analyzing album artwork, or music videos.

Depending on the specific MIR application, researchers or practitioners expect different output values.
Applications that extract audio features typically return audio descriptors, namely values related to the tempo, the key, or the sample rate, among others.
Open source libraries such as \emph{Librosa}\footnote[1]{https://librosa.org/}
and \emph{Essentia}\footnote[2]{https://essentia.upf.edu/index.html} offer methods to extract these values.
Other applications might produce higher-level values, for example, by using machine learning techniques
that estimate the emotion that a track induces, or the music genre of this track.

Among potentially useful input and output values, research has proved Spotify audio features and Last.fm tags to be significant values to characterize music.
Spotify audio features capture high-level information about the music signal,
such as energy, danceability, or valence, among others.
Last.fm tags are text labels that users associate to songs, artists, and albums via the Last.fm social platform.

@@

In particular, this article focuses on predicting Spotify audio features, given a set of Last.fm tags.
By predicting Spotify audio features, we explore the relationship between the subjective perception captured by Last.fm tags, and the concrete musical features that Spotify computes.
This approach might help to identify patterns and hidden correlations between how music is perceived, consumed, and discovered.

@@

% The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
% Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
% This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.

In the remainder of the article, we explain the data gathering and preparation process, as well as the data input formats and varios models.
We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.
\textcolor{red}{UNDER DEVELOPMENT}

\section{Data Sources}

@@

\section{Related Research}

@@

\section{Generating a Dataset}

@@

\section{Experiments}

@@

\begin{figure}[h!]
      \centering
      \includegraphics{images/rmse_by_model_and_limit.png}
      \caption{RMSE mean and standard deviation by model and tags/tokens limit.}
      \label{fig:rmse_by_model_and_limit}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics{images/rmse_by_model_and_feature.png}
      \caption{RMSE mean and standard deviation by model and audio feature.}
      \label{fig:rmse_by_model_and_feature}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics{images/rmse_by_model_and_input.png}
      \caption{RMSE mean and standard deviation by model and input type (tag probablities or tokens).}
      \label{fig:rmse_by_model_and_input}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics[width=1\textwidth]{images/rmse_by_feature_and_input.eps}
      \caption{RMSE mean and standard deviation by audio feature and input type (tag probablities or tokens).}
      \label{fig:rmse_by_feature_and_input}
\end{figure}

\section{Conclusions}

>>>>

The metadata associated to a music piece is available in multiple formats.
Editorial information or lyrics, for example, are mostly available in text format.
The ability to process images or videos might be also required, for example, for analyzing album artwork, or music videos.

Depending on the specific MIR application, researchers and practitioners create models that generate different output values.
Applications that extract audio features typically return audio descriptors, namely values related to the tempo, the key, or the sample rate, to name a few.
Open source libraries such as \emph{Librosa}\footnote[1]{https://librosa.org/}
and \emph{Essentia}\footnote[2]{https://essentia.upf.edu/index.html} offer methods to extract these values.
Other applications might produce more abstract or subjective values, for example, by using machine learning techniques
that estimate the emotion that a track induces, or the music genre of this track.

Among potentially useful input and output values, research has proved \emph{Spotify audio features} and \emph{Last.fm tags} to be significant values used to  characterize music.
Spotify audio features capture high-level information about the music signal and how humans perceive this signal.
Examples of these features are such as energy, danceability, or valence.
Last.fm tags are text labels that users associate to songs, artists, and albums via the Last.fm social platform.

@@

In particular, this article focuses on predicting Spotify audio features, given a set of Last.fm tags.
By predicting Spotify audio features, we explore the relationship between the subjective text descriptions captured by Last.fm tags, and the concrete (but still user-related) musical features that Spotify computes.
This approach might help to identify patterns and hidden correlations between how music is perceived, consumed, and discovered.

@@

The remainder of the article is organized as follows.
Section \ref{sec2} covers the details Last.fm tags and Spotify Audio Features.
Section \ref{sec3} performs a general review of previous research.
Subsequently, section \ref{sec4} and \ref{sec5} explain how the data was gathered and prepared, as well as the conducted experiments.
Finally, in section \ref{sec6} we conclude our study.
% In the remainder of the article, we explain the data gathering and preparation process, as well as the data input formats and varios models.
% We will explore various models for the same track and provide insights on how accurately the prediction can be, by using only Last.fm tags.

\section{Data Sources}\label{sec2}

@@

\section{Related Research}\label{sec3}

@@

\section{Generating a Dataset}\label{sec4}

@@

\section{Experiments}\label{sec5}

@@

\begin{figure}[h!]
      \centering
      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_limit.png}
      \caption{RMSE mean and standard deviation by model and tags/tokens limit.}
      \label{fig:rmse_by_model_and_limit}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_feature.png}
      \caption{RMSE mean and standard deviation by model and audio feature.}
      \label{fig:rmse_by_model_and_feature}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics[width=1\textwidth]{images/rmse_by_model_and_input.png}
      \caption{RMSE mean and standard deviation by model and input type (tag probablities or tokens).}
      \label{fig:rmse_by_model_and_input}
\end{figure}

\begin{figure}[h!]
      \centering
      \includegraphics[width=1\textwidth]{images/rmse_by_feature_and_input.png}
      \caption{RMSE mean and standard deviation by audio feature and input type (tag probablities or tokens).}
      \label{fig:rmse_by_feature_and_input}
\end{figure}

\section{Conclusions}\label{sec6}
